What's hot and what's not:\\Windowed developer topic analysis

### 2009-04-11 lda-paper.tex
submitted once

%\title{Windowed Developer Topic Analysis}
\title{What's hot and what's not:\\Windowing developer topic analysis}

@@

We have devised several techniques for visualizing these topics and
trends.  For all of these techniques if we find trends that have
continuous segments, then we plot those segments as joined
horizontally across time. One technique is the \emph{compact-trend-view},
shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
which displays trends as they occur on the time-based x-axis
(placement along the y-axis is arbitrary).  Another technique is
the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
each trend gets it own distinct y-value , while the x-axis is time;
these topics are then each plotted on their own line across time as
they occur. Our final technique is the \emph{trend-histogram}, shown in Figure
\ref{fig:histogram} where we plot each trend on its own line but stack
up the segments of the trend, much like a histogram.

@@

The \emph{trend-histogram} (Figure \ref{fig:histogram}) superficially
resembles the trend-time-line.  However, in this view the trends are
plotted together by stacking to the left of their row, thus time
information is lost.  The trends are ordered by the number of topics
in the trend.  The trend-histogram shows the count of instances of a
trend and thus indicates which trends occur the most. Unfortunately
due to the large number of topics given allotted space it is often
best to crop off the trends with only one topic, otherwise the tail is
long.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
  \caption{The top part of a trend-histogram of MaxDB 7.500, ordered
    by topic occurrence. X axis determines the number of continuous
    months of a trend. Trends are ranked by the number of topics that
    a trend contains in descending order.}
  \label{fig:histogram}
\end{figure}

@@

We applied our methodology to multiple database systems that we
extracted.  To analyze these extracted repositories we used:
Hiraldo-Grok, an OCaml-based variant of the Grok query language;
Gnuplot, a graph plotting package; lda-c, a LDA package implemented by
Blei et al.~\cite{944937}; and our topic plotter, implemented in
Haskell.

@@

When we examined PostgreSQL, we did not find many trends with two or
more topics with a similarity of $7/10$. Those trends that we did find
were not very large, lasting on 3 months at most. The first and second
largest trends were dedicated to the work of two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the TODO list in HTML, and time string formating topics
relating to timezones.

If we kept the stop words we found that the large trend consisted
mostly of stop words and non-stop words such as \emph{patch},
\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
$1/2$ similarity the largest most the largest most common trend, which
stretched across development, contained these same words
(\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
mentions Dal Zotto, while the third largest trend mentions the
\texttt{[PATCHES]} mailinglist and the names of some patch
contributors.  Other repeating topics refer to portability with Win32,
Central Europe Time (CEST) from email headers, issues with ALTER TABLE
and CVS branch merging (because CVS does not record merges explicitly).

>>>>

%\title{Windowed Developer Topic Analysis}
\title{What's hot and what's not:\\Windowed developer topic analysis}

@@

We have devised several techniques for visualizing these topics and
trends.  For all of these techniques if we find trends that have
continuous segments, then we plot those segments as joined
horizontally across time. One technique is the \emph{compact-trend-view},
shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
which displays trends as they occur on the time-based x-axis
(placement along the y-axis is arbitrary).  Another technique is
the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
each trend gets it own distinct y-value, while the x-axis is time;
these topics are then each plotted on their own line across time as
they occur. Our final technique is the \emph{trend-histogram}, shown in Figure
\ref{fig:histogram} where we plot each trend on its own line but stack
up the segments of the trend, much like a histogram.

@@

The \emph{trend-histogram} (Figure \ref{fig:histogram}) superficially
resembles the trend-time-line.  However, in this view the trends are
plotted together by stacking to the left of their row, thus time
information is lost.  The trends are ordered by the number of topics
in the trend.  The trend-histogram shows the count of instances of a
trend and thus indicates which trends occur the most. Unfortunately
due to the large number of topics, given the allotted space, it is often
best to crop off the trends with only one topic, otherwise the tail is
long.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
  \caption{The top part of a trend-histogram of MaxDB 7.500, ordered
    by topic occurrence. X-axis determines the number of continuous
    months of a trend. Trends are ranked by the number of topics that
    a trend contains in descending order.}
  \label{fig:histogram}
\end{figure}

@@

We applied our methodology to multiple database systems that we
extracted.  To analyze these extracted repositories we used:
Hiraldo-Grok, an OCaml-based variant of the Grok query language;
Gnuplot, a graph plotting package; lda-c, a LDA package implemented by
Blei et al.~\cite{944937}; and our trend plotter, implemented in
Haskell.

@@

When we examined PostgreSQL, we did not find many trends with two or
more topics, using a similarity of $7/10$. Those trends that we did find
were not very large, lasting only 3 months at most. The first and second
largest trends were dedicated to the work of two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the TODO, and time string formating topics
relating to timezones.

If we kept the stop words we found that the large trend consisted
mostly of stop words and non-stop words such as \emph{patch},
\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
$1/2$ similarity the largest most common trend, which
stretched across the entire development, contained these same words
(\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
mentions Dal Zotto, while the third largest trend mentions the
\texttt{[PATCHES]} mailinglist and the names of some patch
contributors.  Other repeating topics refer to portability with Win32,
Central Europe Time (CEST) from email headers, issues with ALTER TABLE
and CVS branch merging, because CVS does not record merges explicitly.
